## install and load libraries
```{r load-library}
library(caret)
library(knitr)
library(corrplot)
library(rpart)
library(rpart.plot)
library(randomForest)
library(gbm)
```

## download data
```{r download-data}
train <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
test  <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))
```

## create partition with the training data set
```{r create-partition}
train_partition  <- createDataPartition(train$classe, p=0.7, list=FALSE)
train_set <- train[train_partition, ]
test_set <- train[-train_partition, ]
```

## remove variables with nearly zero variance
```{r remove-var}
nearly_zero_var <- nearZeroVar(train_set)
train_set <- train_set[, -nearly_zero_var]
test_set  <- test_set[, -nearly_zero_var]
```

## remove variables that are NA
```{r remove-na}
na_var <- sapply(train_set, function(x) mean(is.na(x))) > 0.95
train_set <- train_set[, na_var==FALSE]
test_set  <- test_set[, na_var==FALSE]
```

## remove columns 1 to 5 i.e. identification only variables
```{r remove-col}
train_set <- train_set[, -(1:5)]
test_set  <- test_set[, -(1:5)]
```

## dim training and test set
```{r dim-data}
dim(train_set)
dim(test_set)
```

## prediction using random forest. note: this project will omit using decision tree due to its lower accuracy.
```{r randomforest}
set.seed(343)
control_rf <- trainControl(method="cv", number=3, verboseIter=FALSE)
rf_model <- train(classe ~ ., data=train_set, method="rf", trControl=control_rf)
rf_model$finalModel
```

## assess accuracy of random forest model
```{r rf_accuracy}
rf_predict <- predict(rf_model,newdata = test_set )
rf <- confusionMatrix(test_set$classe,rf_predict)
rf
```

## prediction using generalised boosted model (GBM)
```{r gbm}
set.seed(222)
control_gbm <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
gbm_model  <- train(classe ~ ., data=train_set, method = "gbm",
                    trControl = control_gbm, verbose = FALSE)
gbm_model$finalModel
```

## assess accuracy of gbm model
```{r gbm_accuracy}
gbm_predict <- predict(gbm_model, newdata=test_set)
gbm <- confusionMatrix(gbm_predict, test_set$classe)
gbm
```

## The random forest model is more accurate at ~99% than the GBM model ~98%. Hence, prediction using the random forest model will be used for validation i.e. the predicted results were used to predict the answer for each case.
```{r predict}
predict_quiz <- predict(rf_model,newdata=test)
predict_quiz
```

# the random forest model managed to predict the answer to each question in the quiz with 100% accuracy.